Modeling the Circumnuclear Mediums and Accetion Rates of Quiescent Supermassive Black Holes

### 2015-01-12 ms.tex
editing2

\newcommand\lsim{\mathrel{\rlap{\lower4pt\hbox{\hskip1pt$\sim$}}
    \raise1pt\hbox{$<$}}}
\newcommand\gsim{\mathrel{\rlap{\lower4pt\hbox{\hskip1pt$\sim$}}
    \raise1pt\hbox{$>$}}}
\newcommand{\rs}{r_s}
\newcommand{\rb}{r_b}
\newcommand{\rcirc}{r_{\rm circ}}
\newcommand{\rss}{r_{\rm ss}}
\newcommand{\lrs}{\l_{\rs}}
\newcommand{\lambdars}{\lambda_{\rs}}
\newcommand{\vw}{v_w}

@@

\newcommand{\ke}{\frac{v^2}{2}}
\newcommand{\kew}{\frac{v_w^2}{2}}

@@

\newcommand{\vwO}{v_{w,0}}
\newcommand{\kewO}{\frac{\vwO^2}{2}}
\newcommand{\x}{\frac{r_s}{\rsoi}}
\newcommand{\vwNorm}{\frac{\vwO}{\sigsoi}}
\newcommand{\vwOFH}{v_{w,0,500}}
\newcommand{\vwOFHexp}{\frac{\vwO}{500 \, {\rm km/s}}}

@@

Knowledge of how $\dot{M}$ varies with the SMBH mass, $\Mbh$, and other properties of the
galactic nucleus informs key questions, such as the co-evolution of SMBHs and their host galaxies with cosmic time.  SMBH growth in the low redshift universe appears to be dominated by lower mass black holes (e.g.~\citealt{Heckman+04}).  This is often
attributed to cosmological `down sizing' (e.g, \citealt{Gallo+08}), even though the physical processes involved may be distinct from those operating at higher SMBH masses and accretion rates in quasars.

@@

The term $\propto \vw^2 = \sigma(r)^2+v_{w,0}^2$ in the entropy
equation accounts for external heating sources (e.g.,
\citealt{ShcherbakovWong+:2014a}).  The first term accounts for the
minimum shock heating of the stellar winds due to the random movement
of stars with respect to each other at the characteristic velocity
$\sim \sigma = \sqrt{GM_{\rm enc}/r}$, as is present even for the low
wind velocities $\ll \sigma$ that characterize AGB mass loss in old
stellar populations.  The second term, $v_{w,0}^{2}$, parameterizes
the contribution of additional sources of heating, such as faster main
sequence or Wolf-Rayet winds, millisecond pulsars, supernovae, AGN
feedback, etc.  As discussed in $\S\ref{sec:heating}$, the value of
$v_{w,0}$ is uncertain and will in general depend on the SMBH mass and
the age of the stellar population.

@@

Equations (\ref{eq:drhodt})-(\ref{eq:dsdt}) are solved using a sixth
order finite difference scheme with a third order Runge-Kutta scheme
for time integration and artificial viscosity terms in the velocity
and entropy equations for numerical stability
(\citealt{Brandenburg:2003a}).  We solve the equations for different
choices of $v_{\rm w,0} = 300, 600, 1200$ km s$^{-1}$ spanning a
physically plausible range.  Although we are interested in the
steady-state intflow/outflow structure, the time-dependent equations
are solved to avoid numerical issues that arise near the critical
sonic points, of which there are up to two because of the supersonic
inflow on small radial scales and a possible supersonic outflow on
large scales.

@@

We begin by describing analytic estimates of physical quantities, such as the stagnation radius $\rs$ and the mass accretion rate, the derivation of which are given in Appendix \ref{app:rs}.  Our results are expressed in terms of the SMBH mass $M_{\bullet} = 10^{8}M_{\bullet,8}M_{\odot}$ and the wind heating parameter
(eq.~[\ref{eq:dsdt}])
\begin{eqnarray}
  v_{\rm w} &=& v_{\rm w,0}\left(1+\frac{\sigma^{2}}{v_{\rm w,0}^{2}}\right)^{1/2}  \nonumber \\
  &\approx& v_{\rm w,0}\left(1+0.14 \Mbheight^{0.43}\left[\frac{v_{w,0}}{500\,\rm km\,s^{-1}}\right]^{-1}\right)^{1/2},
\end{eqnarray}
where the final equality makes use of $M_{\bullet}-\sigma$ (eq.~[\ref{eq:Msigma}]).

@@

Continuity of the entropy derivative at the stagnation radius demands that the temperature at the stagnation radius is given by (eq.~[\ref{eq:first_law}])
\begin{align}
T \simeq \frac{\gamma-1}{\gamma}\frac{\mu m_p v_{w}^{2}}{2} \approx 3.7
\times 10^6\ \, v_{500}^2 \,\,{\rm K},
\label{eq:Tanalytic}
\end{align}
where $v_{500} \equiv v_{\rm w}/500$ km s$^{-1}$.

Hydrostatic equilibrium can be used to determine the location of the stagnation radius where $v = 0$ (Appendix \ref{app:rs}, eq.~\ref{eq:rs2main})
\begin{align}
\rs=\frac{G \Mbh}{n \vw(\rs)^2}\left(\frac{9-\Gamma}{2} \frac{\Mstar(\rs)}{\Mbh} +\frac{7}{2}\right),
\label{eq:stag_analytic}
\end{align}
where $n$ is the power-law slope of the density profile at $r = r_{\rm s}$.  If the SMBH dominates the gravitational potential at the stagnation radius then $v_{w,0} \gg \sigma$ and $r_{\rm s} < r_{\rm soi}$ (eq.~[\ref{eq:rsoi}]).  In this case equation (\ref{eq:stag_analytic}) simplifies to
\begin{align}
  &\rs=\frac{7}{2}\frac{G \Mbh}{n v_{\rm w}^{2}} \approx 6
 \, \pc \,\, \Mbheight v_{500}^{-2},
  \label{eq:stag_simple}
\end{align}
where the numerical estimate assumes $n = 1$.  This expression is similar to that obtained by \citet{Volonteri+11} on heuristic grounds (their eq.~6).  In the opposite limit of weak heating ($v_{\rm w,0} \ll \sigma$), the stagnation radius moves to large radii $\gg r_{\rm soi}$ or may not exist at all.  This makes intuitive sense: gas is injected by stars which are gravitationally bound to the black hole, so outflows only occur if the heating rate per gram $\sim v_{\rm w}^{2}$ exceeds the specific gravitational binding energy $\sim \sigma^{2}$.

@@

The SMBH accretion rate is given by the integrated mass loss rate interior to the stagnation radius (eq.~[\ref{eq:q}]), 
\begin{eqnarray}
  \dot{M} &=& \frac{\eta \Menc(\rs)}{\th} \nonumber \\
&\approx&
  \begin{cases}
    2.9 \times 10^{-5} M_{\bullet,8}^{1.76}
    v_{500}^{-3.8}  \eta_{0.02} \Msun \, \pyear& \text{core} \\
    5.3 \times 10^{-5} M_{\bullet,8}^{1.48} 
    v_{500}^{-2.4}  \eta_{0.02} \Msun \, \pyear  & \text{cusp}, 
  \end{cases}
  \label{eq:mdot_analytic}
\end{eqnarray}
where we have made use of equation (\ref{eq:stag_simple}), assuming that $v_{w,0} \gg \sigma$.  The resulting Eddington ratio $\eddr$ is given by 
\begin{eqnarray}
\frac{\dot{M}}{\dot{M}_{\rm edd}} &\approx&
  \begin{cases}
    1.3 \times 10^{-5} M_{\bullet,8}^{0.76}
    v_{500}^{-3.8}  \eta_{0.02}   & \text{core} \\
    2.4 \times 10^{-5} \Mbheight^{0.48} 
    v_{500}^{-2.4}  \eta_{0.02}   & \text{cusp}, 
  \end{cases}
  \label{eq:eddr_analytic}
\end{eqnarray}
%To reproduce the estimated values $\dot{M}/\dot{M}_{\rm edd} \lesssim 10^{-3}$ in low-luminosity AGN or quiescent galaxies, we thus require $v_{w,0} \gtrsim 150M_{\bullet,8}^{0.5}$

@@

The above analytic results neglect the effects of heat conduction near the stagnation radius, an assumption we now check.  The importance of conductive heating can be estimated by comparing its magnitude to the nominal level of ``wind" heating:
\begin{align}
  \left.\frac{\nabla\cdot(\kappa \nabla T)}{q v_{\rm
w}^{2}/2}\right|_{r_{\rm s}} &\sim \frac{2t_{\rm h}\kappa_0
T^{7/2}}{r_{\rm s}^{2}\eta \rho_{\star}(r_{\rm s}) v_{\rm w}^{2}(1+\phi)}
\nonumber \\ &\sim {\rm min}
  \begin{cases}
  10 \eta_{0.02}^{-1}
M_{\bullet,8}^{-0.8} v_{500}^{6.8} &  \text{unsaturated (core)}\\
 10 \eta_{0.02}^{-1}
M_{\bullet,8}^{-0.5} v_{500}^{5.4} &  \text{unsaturated (cusp)}\\
  10 \phi/(2-\Gamma) & \text{saturated},
  \end{cases}
 \label{eq:rhors}
\end{align}
where the second equality makes use of equations (\ref{eq:Tanalytic}), (\ref{eq:stag_simple}), and we have approximated $\nabla^{2} \sim 1/r_{\rm s}^{2}$, $\nabla \sim 1/r_{\rm s}$.  Because the stagnation radius is usually well inside the break radius, we have approximated the stellar density as (eq.~[\ref{eq:rhostar}])
\begin{eqnarray}
  \rho_{\star}(r_{\rm s}) &\simeq& \frac{M_{\bullet}(2-\Gamma)}{4\pi r_{\rm soi}^{3}}\left(\frac{r_{\rm s}}{r_{\rm soi}}\right)^{-1-\Gamma} \nonumber \\
 &\approx& \begin{cases}
    9.4 \times 10^{-19}M_{\bullet,8}^{-1.2}v_{500}^{2.2}\,{\rm g\,cm^{-3}}
    & \text{core} \\
    1.1 \times 10^{-18} M_{\bullet,8}^{-1.5}v_{500}^{3.6}
    \,{\rm g\,cm^{-3}}  & \text{cusp}, 
  \end{cases}
  \label{eq:rhostarrs}
\end{eqnarray}
Even in the saturated case, heat conduction is at most an order unity correction for expected values of the saturation parameter $\phi < 0.1$, justifying its neglect in our analysis.

@@

Radiative cooling can, however, be stabilized by the other sources of cooling, such as the stellar wind source term $\propto -q c_{s}^{2} \propto -qT$ in the entropy equation (eq.~[\ref{eq:dsdt}]), which at the stagnation radius equals $q v_{\rm w}^{2}/2$ (neglecting conduction and radiative cooling).  Thermal stability at the stagnation radius can thus be assessed by comparing the nominal wind heating rate to the rate of radiative cooling $|\dot{q}_{\rm cool}| =|\Lambda(T) n^2|$,
\begin{align}
\left.\frac{q \vw^2/2}{|\dot{q}_{\rm cool}|}\right|_{\rm r{s}} \simeq
  \begin{cases}
   590 \eta_{0.02}^{-1} M_{\bullet,8}^{-0.76}v_{500}^{7.2}  &, \text{core}\\
   210 \eta_{0.02}^{-1} M_{\bullet,8}^{-0.48}v_{500}^{5.8}  &, \text{cusp},     
  \end{cases}
  \label{eq:cooling2}
\end{align}
where we have used equations (\ref{eq:rhors}) and (\ref{eq:rhostarrs}) for the gaseous and stellar densities at the stagnation radius, respectively.  Here $n = \rho/\mu m_p$ and we have approximated the cooling function for $T < 2\times 10^{7}$ K as $\Lambda(T) = 1.1 \times 10^{-22} \left(T/10^6 \text{K}\right)^{-0.7}  $erg cm$^3 $s$^{-1}$ assuming solar metallicity gas (\citealt{Draine:2011a}, Fig. 34.1).  To within a constant of order unity, equation (\ref{eq:cooling2}) also equals the ratio of the gas cooling timescale $t_{\rm cool} \equiv (3\rho kT/2\mu m_p)/\dot{q}_{\rm cool}$ to the free-fall time $t_{\rm ff}$ if one uses the equality 
\begin{align}
\rho(\rs)=\frac{3 q(\rs) \tff(\rs)}{2-\Gamma}.
\label{eq:rhors2}
\end{align}
that results by combining equations~\eqref{eq:mdot_analytic} and~\eqref{eq:mdot_gas}.

Thermal stability ($0.5 q \vw^2/|\dot{q}_{\rm cool}| > 1$; $t_{\rm cool} > t_{\rm ff}$) at the stagnation radius thus requires a critical minimum heating rate
\begin{align}
v_{w,0} > v_{\rm crit} \simeq
  {\rm max}\begin{cases}
   200 \eta_{0.02}^{0.14} M_{\bullet,8}^{0.11}\,{\rm km\,s^{-1}}  &, \text{core}\\
   200 \eta_{0.02}^{0.17} M_{\bullet,8}^{0.08}\,{\rm km\,s^{-1}}   &, \text{cusp}\\
\sigma,     
  \end{cases}
  \label{eq:cooling2}
\end{align}
Our numerical results in $\S\ref{sec:numerical}$ demonstrate that $v_{w} \gtrsim v_{\rm crit} \sim 200-300$ km s$^{-1}$ is generally a condition for stability at all radii, not just at the stagnation radius.  The inclusion of $\sigma$ in equation (\ref{eq:cooling2}) results because the top two lines apply only for $v_{w} > \sigma$ since their derivation assumed the existence of a stagnation radius given by equation (\ref{eq:rs_simple}).

@@

\begin{deluxetable}{lllllll}
\tabletypesize{\scriptsize}
\tabcolsep0.05in\footnotesize
\tablewidth{0pc}
\tablecaption{Cusp Results
\label{tab:sols}}
\tablehead{\colhead{$M_{\bullet,8}$} & \colhead{$v_{w,0}$} &
  \colhead{$\gamma$} & \colhead{$r_b$} &
  \colhead{$\dot{M}/\dot{M}_{\rm Edd}$} & \colhead{$t_{\rm
      cool}/t_{\rm ff} (r_s)$} & \colhead{comments}}
\startdata
0.01 & 300.0 & 0.8 & 100.0 & $ 3.2 \times 10^{ -4 }$ & 4.7 & rb in fact
unresolved \\
0.01 & 600.0 & 0.8 & 100.0 & $ 5.5 \times 10^{ -5 }$ & $ 3.2 \times
10^{ 2 }$ &  rb in fact unresolved\\
0.01 & 1200.0 & 0.8 & 100.0 & $ 9.4 \times 10^{ -6 }$ & $ 1.5 \times
10^{ 4 }$ &  rb in fact unresolved\\
0.1 & 300.0 & 0.8 & 100.0 & $ 2.5 \times 10^{ -3 }$ & 0.29 & energy
cons on outer grid \\
0.1 & 600.0 & 0.8 & 100.0 & $ 1.8 \times 10^{ -4 }$ & 97 &  \\
0.1 & 1200.0 & 0.8 & 100.0 & $ 3.1 \times 10^{ -5 }$ & $ 4.8 \times
10^{ 3 }$ &  energy cons on inner grid\\
1 & 300.0 & 0.8 & 100.0 & $ 1.6 \times 10^{ -2 }$ & $ 5.6 \times 10^{ -2 }$ &  \\
1 & 600.0 & 0.8 & 100.0 & $ 7.3 \times 10^{ -4 }$ & 21 &  \\
1 & 1200.0 & 0.8 & 100.0 & $ 9.9 \times 10^{ -5 }$ & $ 1.5 \times 10^{ 3 }$ &  \\
0.1 & 300.0 & 0.8 & 50.0 & $ 2.0 \times 10^{ -3 }$ & 0.46 &  \\
0.1 & 300.0 & 0.8 & 300.0 & $ 7.2 \times 10^{ -3 }$ & $ 6.7 \times 10^{ -2 }$ &  \\
0.1 & 300.0 & 0.8 & 200.0 & $ 5.3 \times 10^{ -3 }$ & 0.12 &  \\
0.1 & 300.0 & 0.8 & 400.0 & $ 8.9 \times 10^{ -3 }$ & $ 4.7 \times 10^{ -2 }$ &  \\
1 & 300.0 & 0.8 & 50.0 & $ 8.2 \times 10^{ -3 }$ & 0.19 &  \\
1 & 300.0 & 0.8 & 200.0 & $ 3.6 \times 10^{ -2 }$ & $ 1.7 \times 10^{
  -2 }$ &  possible issue w/ conv \\
1 & 300.0 & 0.8 & 300.0 & $ 5.2 \times 10^{ -2 }$ & $ 9.8 \times 10^{
  -3 }$ &  possible issue w/ conv ; rb=400 pc was unsuccessful\\
\enddata
\end{deluxetable}

@@

\begin{deluxetable}{lllllll}
\tabletypesize{\scriptsize}
%\tablecolumns{6}
\tabcolsep0.05in\footnotesize
\tablewidth{0pc}
\tablecaption{Core Results \label{tab:sols}}
\tablehead{\colhead{$M_{\bullet,8}$} & \colhead{$v_{w,0}$} &
\colhead{$\gamma$} & \colhead{$r_b$} & \colhead{$\dot{M}/\dot{M}_{\rm
Edd}$}  & \colhead{$t_{\rm cool}/t_{\rm ff} (r_s)$} & \colhead{comments}}
\startdata
0.01 & 600.0 & 0.1 & 100.0 & $ 3.4 \times 10^{ -5 }$ & $ 1.2 \times 10^{ 2 }$ &  \\
0.01 & 1200.0 & 0.1 & 100.0 & $ 1.5 \times 10^{ -6 }$ & $ 4.1 \times 10^{ 4 }$ &  \\
0.1 & 1200.0 & 0.1 & 100.0 & $ 9.5 \times 10^{ -6 }$ & $ 5.7 \times 10^{ 3 }$ &  \\
1 & 1200.0 & 0.1 & 100.0 & $ 6.0 \times 10^{ -5 }$ & $ 8.6 \times 10^{ 2 }$ &  \\
1 & 1200.0 & 0.1 & 300.0 & $ 1.9 \times 10^{ -4 }$ & $ 1.3 \times 10^{ 2 }$ &  \\
\enddata
\end{deluxetable}

This section describes the results of our numerical calculations,
which we use to assess the validity of the analytic expressions
provided in the previous section.  Numerical solutions are calculated
for each galaxy for three different values of the stellar wind heating
parameter $v_{w,0}$ = 300, 600, 1200 km s$^{-1}$.

>>>>

\newcommand\lsim{\mathrel{\rlap{\lower4pt\hbox{\hskip1pt$\sim$}}
    \raise1pt\hbox{$<$}}}
\newcommand\gsim{\mathrel{\rlap{\lower4pt\hbox{\hskip1pt$\sim$}}
    \raise1pt\hbox{$>$}}}
\newcommand{\rs}{r_s}
\newcommand{\rb}{r_b}
\newcommand{\rcirc}{r_{\rm circ}}
\newcommand{\rss}{r_{\rm ss}}
\newcommand{\lrs}{\l_{\rs}}
\newcommand{\lambdars}{\lambda_{\rs}}
\newcommand{\vw}{\tilde{v}_{w}}

@@

\newcommand{\ke}{\frac{v^2}{2}}
\newcommand{\kew}{\frac{\tilde{v}_{w}^2}{2}}

@@

\newcommand{\vwO}{v_{w}}
\newcommand{\kewO}{\frac{\vwO^2}{2}}
\newcommand{\x}{\frac{r_s}{\rsoi}}
\newcommand{\vwNorm}{\frac{\vwO}{\sigsoi}}
\newcommand{\vwOFH}{v_{500}}
\newcommand{\vwOFHexp}{\frac{\vwO}{500 \, {\rm km/s}}}

@@

Knowledge of how $\dot{M}$ varies with the SMBH mass, $\Mbh$, and other properties of the
galactic nucleus informs key questions, such as the co-evolution of SMBHs and their host galaxies with cosmic time.  SMBH growth in the low redshift universe appears to be dominated by lower mass black holes (e.g.~\citealt{Heckman+04}).  This is often
attributed to general trend  of cosmological `down-sizing' (e.g, \citealt{Gallo+08}).  However, the physical processes involved may be distinct from those operating at higher SMBH masses and accretion rates in quasars.

@@

The term $\propto \vw^2 = \sigma(r)^2+v_{w}^2$ in the entropy
equation accounts for external heating sources (e.g.,
\citealt{ShcherbakovWong+:2014a}).  The first term accounts for the
minimum shock heating of the stellar winds due to the random movement
of stars with respect to each other at the characteristic velocity
$\sim \sigma = \sqrt{GM_{\rm enc}/r}$, as is present even for the low
wind velocities $\ll \sigma$ that characterize AGB mass loss in old
stellar populations.  The second term, $v_{w}^{2}$, parameterizes
the contribution of additional sources of heating, such as faster main
sequence or Wolf-Rayet winds, millisecond pulsars, supernovae, AGN
feedback, etc.  As discussed in $\S\ref{sec:heating}$, the value of
$v_{w}$ is uncertain and will in general depend on the SMBH mass and
the age of the stellar population.

@@

Equations (\ref{eq:drhodt})-(\ref{eq:dsdt}) are solved using a sixth
order finite difference scheme with a third order Runge-Kutta scheme
for time integration and artificial viscosity terms in the velocity
and entropy equations for numerical stability
(\citealt{Brandenburg:2003a}).  We solve the equations for different
choices of $v_{w} = 300, 600, 1200$ km s$^{-1}$ spanning a
physically plausible range.  Although we are interested in the
steady-state intflow/outflow structure, the time-dependent equations
are solved to avoid numerical issues that arise near the critical
sonic points, of which there are up to two because of the supersonic
inflow on small radial scales and a possible supersonic outflow on
large scales.

@@

We begin by describing analytic estimates of physical quantities, such as the stagnation radius $\rs$ and the mass accretion rate, the derivation of which are given in Appendix \ref{app:rs}.  Our results are expressed in terms of the SMBH mass $M_{\bullet}$, heating parameter $\tilde{v}_{w}$, and wind mass loss parameter $\eta$.

@@

Continuity of the entropy derivative at the stagnation radius demands that the temperature at the stagnation radius is given by (eq.~[\ref{eq:first_law}])
\begin{align}
T \simeq \frac{\gamma-1}{\gamma}\frac{\mu m_p \tilde{v}_{w}^{2}}{2} \approx 3.7
\times 10^6\ \, v_{500}^2 \,\,{\rm K},
\label{eq:Tanalytic}
\end{align}
where $v_{500} \equiv \tilde{v}_{w}/500$ km s$^{-1}$.

Hydrostatic equilibrium can be used to determine the location of the stagnation radius where $v = 0$ (Appendix \ref{app:rs}, eq.~\ref{eq:rs2main})
\begin{align}
\rs=\frac{G \Mbh}{n \vw(\rs)^2}\left(\frac{9-\Gamma}{2} \frac{\Mstar(\rs)}{\Mbh} +\frac{7}{2}\right),
\label{eq:stag_analytic}
\end{align}
where $n$ is the power-law slope of the density profile at $r = r_{\rm s}$.  If the SMBH dominates the gravitational potential at the stagnation radius then $v_{w} \gg \sigma$ and $r_{\rm s} < r_{\rm soi}$ (eq.~[\ref{eq:rsoi}]).  In this case equation (\ref{eq:stag_analytic}) simplifies to
\begin{align}
  &\rs=\frac{7}{2}\frac{G \Mbh}{n \tilde{v}_{w}^{2}} \approx 6
 \, \pc \,\, \Mbheight v_{500}^{-2},
  \label{eq:stag_simple}
\end{align}
where the numerical estimate assumes $n = 1$.  This expression is similar to that obtained by \citet{Volonteri+11} on heuristic grounds (their eq.~6).  In the opposite limit of weak heating ($v_{w} \ll \sigma$), the stagnation radius moves to large radii $\gg r_{\rm soi}$ or may not exist at all.  This makes intuitive sense: gas is injected by stars which are gravitationally bound to the black hole, so outflows only occur if the heating rate per gram $\sim \tilde{v}_{w}^{2}$ exceeds the specific gravitational binding energy $\sim \sigma^{2}$.

@@

The SMBH accretion rate is given by the integrated mass loss rate interior to the stagnation radius (eq.~[\ref{eq:q}]), 
\begin{eqnarray}
  \dot{M} &=& \frac{\eta \Menc(\rs)}{\th} \nonumber \\
&\approx&
  \begin{cases}
    2.9 \times 10^{-5} M_{\bullet,8}^{1.76}
    v_{500}^{-3.8}  \eta_{0.02} \Msun \, \pyear& \text{core} \\
    5.3 \times 10^{-5} M_{\bullet,8}^{1.48} 
    v_{500}^{-2.4}  \eta_{0.02} \Msun \, \pyear  & \text{cusp}, 
  \end{cases}
  \label{eq:mdot_analytic}
\end{eqnarray}
where we have made use of equation (\ref{eq:stag_simple}), assuming that $v_{w} \gg \sigma$.  The resulting Eddington ratio $\eddr$ is given by 
\begin{eqnarray}
\frac{\dot{M}}{\dot{M}_{\rm edd}} &\approx&
  \begin{cases}
    1.3 \times 10^{-5} M_{\bullet,8}^{0.76}
    v_{500}^{-3.8}  \eta_{0.02}   & \text{core} \\
    2.4 \times 10^{-5} \Mbheight^{0.48} 
    v_{500}^{-2.4}  \eta_{0.02}   & \text{cusp}, 
  \end{cases}
  \label{eq:eddr_analytic}
\end{eqnarray}
%To reproduce the estimated values $\dot{M}/\dot{M}_{\rm edd} \lesssim 10^{-3}$ in low-luminosity AGN or quiescent galaxies, we thus require $v_{w} \gtrsim 150M_{\bullet,8}^{0.5}$

@@

The above analytic results neglect the effects of heat conduction near the stagnation radius, an assumption we now check.  The importance of conductive heating can be estimated by comparing its magnitude to the nominal level of ``wind" heating:
\begin{align}
  \left.\frac{\nabla\cdot(\kappa \nabla T)}{q v_{\rm
w}^{2}/2}\right|_{r_{\rm s}} &\sim \frac{2t_{\rm h}\kappa_0
T^{7/2}}{r_{\rm s}^{2}\eta \rho_{\star}(r_{\rm s}) \tilde{v}_{w}^{2}(1+\phi)}
\nonumber \\ &\sim {\rm min}
  \begin{cases}
  10 \eta_{0.02}^{-1}
M_{\bullet,8}^{-0.8} v_{500}^{6.8} &  \text{unsaturated (core)}\\
 10 \eta_{0.02}^{-1}
M_{\bullet,8}^{-0.5} v_{500}^{5.4} &  \text{unsaturated (cusp)}\\
  10 \phi/(2-\Gamma) & \text{saturated},
  \end{cases}
 \label{eq:rhors}
\end{align}
where the second equality makes use of equations (\ref{eq:Tanalytic}), (\ref{eq:stag_simple}), and we have approximated $\nabla^{2} \sim 1/r_{\rm s}^{2}$, $\nabla \sim 1/r_{\rm s}$.  Because the stagnation radius is usually well inside the break radius, we have approximated the stellar density as (eq.~[\ref{eq:rhostar}])
\begin{eqnarray}
  \rho_{\star}(r_{\rm s}) &\simeq& \frac{M_{\bullet}(2-\Gamma)}{4\pi r_{\rm soi}^{3}}\left(\frac{r_{\rm s}}{r_{\rm soi}}\right)^{-1-\Gamma} \nonumber \\
 &\approx& \begin{cases}
    9.4 \times 10^{-19}M_{\bullet,8}^{-1.2}v_{500}^{2.2}\,{\rm g\,cm^{-3}}
    & \text{core} \\
    1.1 \times 10^{-18} M_{\bullet,8}^{-1.5}v_{500}^{3.6}
    \,{\rm g\,cm^{-3}}  & \text{cusp}, 
  \end{cases}
  \label{eq:rhostarrs}
\end{eqnarray}
Even in the saturated case, heat conduction is at most an order unity correction for expected values of the saturation parameter $\phi < 0.1$, justifying its neglect in our analysis.

@@

Radiative cooling can, however, be stabilized by the other sources of cooling, such as the stellar wind source term $\propto -q c_{s}^{2} \propto -qT$ in the entropy equation (eq.~[\ref{eq:dsdt}]), which at the stagnation radius equals $q \tilde{v}_{w}^{2}/2$ (neglecting conduction and radiative cooling).  Thermal stability at the stagnation radius can thus be assessed by comparing the nominal wind heating rate to the rate of radiative cooling $|\dot{q}_{\rm cool}| =|\Lambda(T) n^2|$,
\begin{align}
\left.\frac{q \vw^2/2}{|\dot{q}_{\rm cool}|}\right|_{\rm r{s}} \simeq
  \begin{cases}
   590 \eta_{0.02}^{-1} M_{\bullet,8}^{-0.76}v_{500}^{7.2}  &, \text{core}\\
   210 \eta_{0.02}^{-1} M_{\bullet,8}^{-0.48}v_{500}^{5.8}  &, \text{cusp},     
  \end{cases}
  \label{eq:cooling2}
\end{align}
where we have used equations (\ref{eq:rhors}) and (\ref{eq:rhostarrs}) for the gaseous and stellar densities at the stagnation radius, respectively.  Here $n = \rho/\mu m_p$ and we have approximated the cooling function for $T < 2\times 10^{7}$ K as $\Lambda(T) = 1.1 \times 10^{-22} \left(T/10^6 \text{K}\right)^{-0.7}  $erg cm$^3 $s$^{-1}$ assuming solar metallicity gas (\citealt{Draine:2011a}, Fig. 34.1).  To within a constant of order unity, equation (\ref{eq:cooling2}) also equals the ratio of the gas cooling timescale $t_{\rm cool} \equiv (3\rho kT/2\mu m_p)/\dot{q}_{\rm cool}$ to the free-fall time $t_{\rm ff}$ if one uses the equality 
\begin{align}
\rho(\rs)=\frac{3 q(\rs) \tff(\rs)}{2-\Gamma}.
\label{eq:rhors2}
\end{align}
that results by combining equations~\eqref{eq:mdot_analytic} and~\eqref{eq:mdot_gas}.

Thermal stability ($0.5 q \vw^2/|\dot{q}_{\rm cool}| > 1$; $t_{\rm cool} > t_{\rm ff}$) at the stagnation radius thus requires a critical minimum heating rate
\begin{align}
v_{w} > v_{\rm crit} \simeq
  {\rm max}\begin{cases}
   200 \eta_{0.02}^{0.14} M_{\bullet,8}^{0.11}\,{\rm km\,s^{-1}}  &, \text{core}\\
   200 \eta_{0.02}^{0.17} M_{\bullet,8}^{0.08}\,{\rm km\,s^{-1}}   &, \text{cusp}\\
\sigma,     
  \end{cases}
  \label{eq:cooling2}
\end{align}
Our numerical results in $\S\ref{sec:numerical}$ demonstrate that $\tilde{v}_{w} \gtrsim v_{\rm crit} \sim 200-300$ km s$^{-1}$ is generally a condition for stability at all radii, not just at the stagnation radius.  The inclusion of $\sigma$ in equation (\ref{eq:cooling2}) results because the top two lines apply only for $\tilde{v}_{w} > \sigma$ since their derivation assumed the existence of a stagnation radius given by equation (\ref{eq:rs_simple}).

@@

\begin{deluxetable}{lllllll}
\tabletypesize{\scriptsize}
\tabcolsep0.05in\footnotesize
\tablewidth{0pc}
\tablecaption{Cusp Results
\label{tab:sols}}
\tablehead{\colhead{$M_{\bullet,8}$} & \colhead{$v_{w}$} &
  \colhead{$\gamma$} & \colhead{$r_b$} &
  \colhead{$\dot{M}/\dot{M}_{\rm Edd}$} & \colhead{$t_{\rm
      cool}/t_{\rm ff} (r_s)$} & \colhead{comments}}
\startdata
0.01 & 300.0 & 0.8 & 100.0 & $ 3.2 \times 10^{ -4 }$ & 4.7 & rb in fact
unresolved \\
0.01 & 600.0 & 0.8 & 100.0 & $ 5.5 \times 10^{ -5 }$ & $ 3.2 \times
10^{ 2 }$ &  rb in fact unresolved\\
0.01 & 1200.0 & 0.8 & 100.0 & $ 9.4 \times 10^{ -6 }$ & $ 1.5 \times
10^{ 4 }$ &  rb in fact unresolved\\
0.1 & 300.0 & 0.8 & 100.0 & $ 2.5 \times 10^{ -3 }$ & 0.29 & energy
cons on outer grid \\
0.1 & 600.0 & 0.8 & 100.0 & $ 1.8 \times 10^{ -4 }$ & 97 &  \\
0.1 & 1200.0 & 0.8 & 100.0 & $ 3.1 \times 10^{ -5 }$ & $ 4.8 \times
10^{ 3 }$ &  energy cons on inner grid\\
1 & 300.0 & 0.8 & 100.0 & $ 1.6 \times 10^{ -2 }$ & $ 5.6 \times 10^{ -2 }$ &  \\
1 & 600.0 & 0.8 & 100.0 & $ 7.3 \times 10^{ -4 }$ & 21 &  \\
1 & 1200.0 & 0.8 & 100.0 & $ 9.9 \times 10^{ -5 }$ & $ 1.5 \times 10^{ 3 }$ &  \\
0.1 & 300.0 & 0.8 & 50.0 & $ 2.0 \times 10^{ -3 }$ & 0.46 &  \\
0.1 & 300.0 & 0.8 & 300.0 & $ 7.2 \times 10^{ -3 }$ & $ 6.7 \times 10^{ -2 }$ &  \\
0.1 & 300.0 & 0.8 & 200.0 & $ 5.3 \times 10^{ -3 }$ & 0.12 &  \\
0.1 & 300.0 & 0.8 & 400.0 & $ 8.9 \times 10^{ -3 }$ & $ 4.7 \times 10^{ -2 }$ &  \\
1 & 300.0 & 0.8 & 50.0 & $ 8.2 \times 10^{ -3 }$ & 0.19 &  \\
1 & 300.0 & 0.8 & 200.0 & $ 3.6 \times 10^{ -2 }$ & $ 1.7 \times 10^{
  -2 }$ &  possible issue w/ conv \\
1 & 300.0 & 0.8 & 300.0 & $ 5.2 \times 10^{ -2 }$ & $ 9.8 \times 10^{
  -3 }$ &  possible issue w/ conv ; rb=400 pc was unsuccessful\\
\enddata
\end{deluxetable}

@@

\begin{deluxetable}{lllllll}
\tabletypesize{\scriptsize}
%\tablecolumns{6}
\tabcolsep0.05in\footnotesize
\tablewidth{0pc}
\tablecaption{Core Results \label{tab:sols}}
\tablehead{\colhead{$M_{\bullet,8}$} & \colhead{$v_{w}$} &
\colhead{$\gamma$} & \colhead{$r_b$} & \colhead{$\dot{M}/\dot{M}_{\rm
Edd}$}  & \colhead{$t_{\rm cool}/t_{\rm ff} (r_s)$} & \colhead{comments}}
\startdata
0.01 & 600.0 & 0.1 & 100.0 & $ 3.4 \times 10^{ -5 }$ & $ 1.2 \times 10^{ 2 }$ &  \\
0.01 & 1200.0 & 0.1 & 100.0 & $ 1.5 \times 10^{ -6 }$ & $ 4.1 \times 10^{ 4 }$ &  \\
0.1 & 1200.0 & 0.1 & 100.0 & $ 9.5 \times 10^{ -6 }$ & $ 5.7 \times 10^{ 3 }$ &  \\
1 & 1200.0 & 0.1 & 100.0 & $ 6.0 \times 10^{ -5 }$ & $ 8.6 \times 10^{ 2 }$ &  \\
1 & 1200.0 & 0.1 & 300.0 & $ 1.9 \times 10^{ -4 }$ & $ 1.3 \times 10^{ 2 }$ &  \\
\enddata
\end{deluxetable}

This section describes the results of our numerical calculations,
which we use to assess the validity of the analytic expressions
provided in the previous section.  Numerical solutions are calculated
for each galaxy for three different values of the stellar wind heating
parameter $v_{w}$ = 300, 600, 1200 km s$^{-1}$.  In cases when the stagnation radius become very large and was difficult to resolve on the grid, we conclude that such solutions will be thermally unstable and terminated their calculations (these are marked as `thermally unstable' in the tables.
